Dish: Distribution-oblivious Shell Scripting

### 2019-11-23 pldi/draft.tex
clean up in 6

\begin{table*}[t]
\center
\footnotesize
% \setlength\tabcolsep{3pt}
\caption{
  \footnotesize{
    \textbf{Summary of Micro-benchmarks}.
    The micro-benchmarks are small pipelines (\ie ones developed on-the-fly) drawn from various sources and applied to large datasets.
  }
}
\begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} llllll}
\toprule
  Script                 ~&~ Structure                    & Input (MB)& Time (Seq)    & Script Size (LoC)             & Highlights                                        \\
\midrule
% https://github.com/andromeda/sdsh/blob/master/scripts/grep.sh
  \tti{min-grep}         ~&~$3\times$\tsta                & 1000      &  78m 36.77s   &   30\qquad 134\qquad  1304    & complex NFA regex                                 \\
% https://github.com/andromeda/sdsh/blob/master/scripts/minimal5.sh
  \tti{min-sort}         ~&~~$\tsta, \tpur$               & 10000     &  23m 19.85s   &   33\qquad 129\qquad  1209    & \tti{sort}ing                                     \\
% https://github.com/andromeda/sdsh/blob/master/scripts/topn.sh
  \tti{wf}               ~&~~$3\times\tsta, 3\times\tpur$ & 1000      &   2m 11.18s   &   53\qquad 181\qquad  1621    & double \tti{sort}, \tti{uniq} reduction           \\
% https://github.com/andromeda/sdsh/blob/master/scripts/topn.sh
  \tti{top-n}            ~&~~$2\times\tsta, 4\times\tpur$ & 1000      &   6m 53.55s   &   53\qquad 181\qquad  1621    & double \tti{sort}, \tti{uniq} reduction           \\
% https://github.com/andromeda/sdsh/blob/master/scripts/ngrams.sh
  \tti{bi-gram}          ~&~~$2\times\tsta, 4\times\tpur$ & \todo{X}  &               &   X \qquad  X \qquad          & stream shifting and merging                       \\
% https://github.com/andromeda/sdsh/blob/master/scripts/spell.sh
  \tti{spell}            ~&~~$4\times\tsta, 3\times\tpur$ & 1000      &   8m 13.42s   &  37 \qquad 257\qquad  2417    & comparisons (\tti{comm})                          \\
% https://github.com/andromeda/sdsh/blob/master/scripts/diff.sh
  \tti{diff}             ~&~~$4\times\tsta, 3\times\tpur$ & \todo{X}  &               &   X \qquad  X \qquad          & shuffling and non-distributable \tti{diff}ing     \\
\bottomrule
\end{tabular*}
\label{tab:eval}
\end{table*}

At a high level, we are interested in understanding whether \sys can indeed scale pipelines out automatically and correctly.
To achieve this, we use a combination of micro- and macro-benchmarks, most of which were collected from prior literature and online repositories.
Micro-benchmarks~\sx{micro} are one-off, simple one-line pipelines that take a few seconds to write (usually interactively);
  such pipelines are usually composed interactively to solve a task at hand, test a hypothesis, or ``smoke out bugs''~\cite{bentley1986literate}---but, with \sys, applied to very large data sets.
These pipelines are taken from real use cases (often from the literature~\cite{bentley1985spelling, bentley1985spelling}), and, precisely due to their small size, have the benefit of highlighting stress on a handful of stages.

Three complex macro-benchmarks (\S\ref{macro1}--\ref{macro3}) involve third-party commands and handle realistic workloads by today's standards.
These scripts are about an order of magnitude larger than the micro-benchmarks, but this perceived lack of complexity is deceiving:
  as shown below, if written in conventional languages, these pipelines would correspond to programs on the order of hundreds of lines of code.

Several results are worth highlighting.
First and foremost, programs see a significant speedup, often more $10\times$ over sequential execution  while returning identical results (on very large datasets).
\sys's transformation and planning phase take around XX--XXms, negligible (XX\%) for short-running pipelines and virtually non-existent for long-running ones.
As \sys rewrites to highly-optimized shell primitives rather than using a managed language runtime, it always has a COST~\cite{mcsherryscalability} of 2.
Finally, \sys preserves the productivity of the shell:
  in one case, the equivalent program for a small part of the pipeline requires over a 150 lines of code in combined Java and Bash to run distributed on top of Hadoop;
  in another, the program expressed by the pipeline was the subject of a semester project in a graduate course on distributed systems~\cite{blinded} where student implementations ranged between 1--3\textbf{K} lines of code.

@@

\subsection{Microbenchmarks: Shell One-liners}

@@

In terms of benchmarks, \ttt{grep} is a short script centered around an expensive \sta command.
% a DFA-based regular expression matching phase;
  while \ttt{grep} defaults to a Thompson NFA, this particular expression makes use of DFA-based backtracking patterns resulting in high runtime overhead.
The \ttt{sort} pipeline is equally short, but its central operation is a command in \pur.
The next two benchmarks, \ttt{wf} and \ttt{top-n}, are based on McIlroy's now classic word-counting program~\cite{bentley1986literate};
  they use sorting, rather than tabulation, to identify high-frequency terms in a corpus.
The \ttt{bi-gram} pipeline calculates n-grams of base two;
  it makes clever use of \ttt{tail} to shift a stream by one word and \ttt{paste} to fuse (zip) two streams together.
The next pipeline is based on the original \ttt{spell} program developed by Johnson~\cite{bentley1985spelling}---another \unix classic:
  after some preprocessing, it makes clever use of \ttt{comm} to report words not in the dictionary.

% TODO \nv{Need to change terminology in plots---files should be (attempted) parallelism, name benchmark, cat should be merge.}
Fig.~\ref{fig:microbenchmarks} presents the speedup gained by \sys as a function of the level of parallelism (1--200).
% we experiment with a $200\times$ level of parallelism to see potential overheads
Each one of these plots reports on three different configurations:
  (i) distributed, which captures only the runtime of the distributed computation,
  (ii) compile + distributed, which adds the \sys' analysis and generation overhead, and
  (iii) cat + compile + distributed, which adds a final merger  gathering the results at the end of the pipeline.
The results show significant speedups, ranging between \todo{7--70$\times$}, depending on the distributability characteristics of individual pipelines.

>>>>

\begin{table*}[t]
\center
\footnotesize
% \setlength\tabcolsep{3pt}
\caption{
  \footnotesize{
    \textbf{Summary of Micro-benchmarks}.
    The micro-benchmarks are small pipelines (\ie ones developed on-the-fly) drawn from various sources and applied to large datasets.
  }
}
\begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} llllll}
\toprule
  Script                 ~&~ Structure                    & Input (MB)& Time (Seq)    & Script Size (LoC)             & Highlights                                        \\
\midrule
% https://github.com/andromeda/sdsh/blob/master/scripts/grep.sh
  \tti{grep}         ~&~$3\times$\tsta                & 1000      &  78m 36.77s   &   30\qquad 134\qquad  1304    & complex NFA regex                                 \\
% https://github.com/andromeda/sdsh/blob/master/scripts/minimal5.sh
  \tti{sort}         ~&~~$\tsta, \tpur$               & 10000     &  23m 19.85s   &   33\qquad 129\qquad  1209    & \tti{sort}ing                                     \\
% https://github.com/andromeda/sdsh/blob/master/scripts/topn.sh
  \tti{wf}               ~&~~$3\times\tsta, 3\times\tpur$ & 1000      &   2m 11.18s   &   53\qquad 181\qquad  1621    & double \tti{sort}, \tti{uniq} reduction           \\
% https://github.com/andromeda/sdsh/blob/master/scripts/topn.sh
  \tti{top-n}            ~&~~$2\times\tsta, 4\times\tpur$ & 1000      &   6m 53.55s   &   53\qquad 181\qquad  1621    & double \tti{sort}, \tti{uniq} reduction           \\
% https://github.com/andromeda/sdsh/blob/master/scripts/ngrams.sh
  \tr{\tti{bi-gram}          ~&~~$2\times\tsta, 4\times\tpur$ & \todo{X}  &               &   X \qquad  X \qquad          & stream shifting and merging                       \\}
% https://github.com/andromeda/sdsh/blob/master/scripts/spell.sh
  \tti{spell}            ~&~~$4\times\tsta, 3\times\tpur$ & 1000      &   8m 13.42s   &  37 \qquad 257\qquad  2417    & comparisons (\tti{comm})                          \\
% https://github.com/andromeda/sdsh/blob/master/scripts/diff.sh
  \tr{\tti{diff}             ~&~~$4\times\tsta, 3\times\tpur$ & \todo{X}  &               &   X \qquad  X \qquad          & shuffling and non-distributable \tti{diff}ing     \\}
  \tti{shortest-scripts} ~&~~$5\times\tsta, 2\times\tpur$ & 8.6       &   2m 50.03s   &  61 \qquad 253\qquad  2413    & \todo{x}                          \\
\bottomrule
\end{tabular*}
\label{tab:eval}
\end{table*}

At a high level, we are interested in understanding whether \sys can indeed scale pipelines out automatically and correctly.
To achieve this, we use a combination of micro-benchmarks and one macro-benchmark, most of which were collected from prior literature and online repositories.
Micro-benchmarks~\sx{micro} are one-off, simple one-line pipelines that take a few seconds to write;
  such pipelines are usually composed interactively to solve a task at hand, test a hypothesis, or ``smoke out bugs''~\cite{bentley1986literate}---but, with \sys, applied to very large data sets.
These pipelines are taken from real use cases (often from the literature~\cite{bentley1985spelling}), and, precisely due to their small size, have the benefit of highlighting stress on a handful of stages.

The complex macro-benchmark (\S\ref{macro1}) is designed to handle realistic workloads by today's standards.
It is about an order of magnitude larger than the micro-benchmarks, but this perceived lack of complexity is deceiving:
  as shown below, if written in conventional languages, this pipeline would correspond to a program on the order of hundreds of lines of code.

Several results are worth highlighting.
First and foremost, programs see a significant speedup, often more $10\times$ over sequential execution  while returning identical results (on very large datasets).
\sys's transformation and planning phase take around 50--300ms, negligible ($<$5\%) for short-running pipelines and virtually non-existent for long-running ones.
As \sys rewrites to highly-optimized shell primitives rather than using a managed language runtime, it always has a COST~\cite{mcsherryscalability} of 2.
Finally, \sys preserves the productivity of the shell:
  the equivalent program for a small part of the macro-benchmark pipeline requires over a 150 lines of code in combined Java and Bash to run distributed on top of Hadoop.
  %% in another, the program expressed by the pipeline was the subject of a semester project in a graduate course on distributed systems~\cite{blinded} where student implementations ranged between 1--3\textbf{K} lines of code.

@@

\subsection{Microbenchmarks: Shell One-liners}
\label{micro}

@@

In terms of benchmarks, \ttt{grep} is a short script centered around an expensive \sta command.
% a DFA-based regular expression matching phase;
  while \ttt{grep} defaults to a Thompson NFA, this particular expression makes use of DFA-based backtracking patterns resulting in high runtime overhead.
The \ttt{sort} pipeline is equally short, but its central operation is a command in \pur.
The next two benchmarks, \ttt{wf} and \ttt{top-n}, are based on McIlroy's now classic word-counting program~\cite{bentley1986literate};
  they use sorting, rather than tabulation, to identify high-frequency terms in a corpus.
\tr{The \ttt{bi-gram} pipeline calculates n-grams of base two;
  it makes clever use of \ttt{tail} to shift a stream by one word and \ttt{paste} to fuse (zip) two streams together.}
The next pipeline is based on the original \ttt{spell} program developed by Johnson~\cite{bentley1985spelling}---another \unix classic:
  after some preprocessing, it makes clever use of \ttt{comm} to report words not in the dictionary.

% TODO \nv{Need to change terminology in plots---files should be (attempted) parallelism, name benchmark, cat should be merge.}
Fig.~\ref{fig:microbenchmarks} presents the speedup gained by \sys as a function of the level of parallelism (1--200).
% we experiment with a $200\times$ level of parallelism to see potential overheads
Each one of these plots reports on three different configurations:
  (i) distributed, which captures only the runtime of the distributed computation,
  (ii) compile + distributed, which adds the \sys' analysis and generation overhead, and
  (iii) cat + compile + distributed, which adds a final merger  gathering the results at the end of the pipeline.
The results show significant speedups, ranging between 4--70$\times$, depending on the distributability characteristics of individual pipelines.
